---
title: "ADSD"
author: "Simon Schwab"
date: "30 May 2017"
output: html_document
---

## Install packages
```{r Install packages}
# install.packages("rmarkdown")
# install.packages("testit")
# install.packages("Hmisc")
# install.packages("nortest")
# install.packages("multdyn")
# install.packages("perm")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
```{r Load libraries, message=FALSE, warning=TRUE}
library(testit)
library(reshape2)
library(ggplot2)
library(cowplot)
library(Hmisc)
library(nortest)
library(multdyn)
library(perm)
```

## Main variables
```{r Main variables}
N = 62
N_ROIS = c(14, 200)
PATH_HOME = "/home/simon"
PATH_ROIS200 = file.path(PATH_HOME, 'Drive', '200RoisN62')
PATH = file.path(PATH_HOME, 'workspace', 'ADSD-nbook')
```

## Subject infos
```{r Subjects infos}
infos = read.table(file.path(PATH, 'ADSD_info.txt'), header = T)
rois = read.table(file.path(PATH, 'ROIs_14.txt'), header = T)
# Excluded subjects
infos$included = rep(TRUE,N)
# DVARS; more than 20% volumes affected:
infos$included[match("ADSD_13", infos$id)] = FALSE
infos$included[match("ADSD_36", infos$id)] = FALSE
# High E-vars:
infos$included[match("ADSD_34", infos$id)] = FALSE
# DVARS below 40%:
infos$included[match("ADSD_15", infos$id)] = FALSE
infos$included[match("ADSD_42", infos$id)] = FALSE
infos$included[match("ADSD_43", infos$id)] = FALSE

infos$included[match("2026", infos$id)] = FALSE # no mri data

infos$isHC = as.logical(infos$isHC)
infos$isAD = as.logical(infos$isAD)
infos$isSD = as.logical(infos$isSD)

infos$group=rep(NA,N)
infos$group[infos$isSD] = 'SD'
infos$group[infos$isAD] = 'AD'
infos$group[infos$isHC] = 'HC'
infos$group = as.factor(infos$group)

assert(nrow(infos) == N)
N_SD  = sum(infos$isSD & infos$included)
N_AD  = sum(infos$isAD & infos$included)
N_HC  = sum(infos$isHC & infos$included)

# DVARS
dvars=read.table(file.path(PATH, 'DVARS.txt'), header = T)
assert(nrow(dvars) == nrow(infos))
infos = cbind(infos, dvars)
rm(dvars)

print(infos[1:10,])
```
## Demographics
```{r Demographics}
d.demog = data.frame(Group=as.factor(c('SD', 'AD', 'HC')),
                     N_Total=c(sum(infos$isSD), sum(infos$isAD), sum(infos$isHC)),
                     N_included=c(sum(infos$isSD & infos$included), sum(infos$isAD & infos$included),
                                  sum(infos$isHC & infos$included)),
                     age_mean=round(c(mean(infos$age[infos$isSD & infos$included]),
                                 mean(infos$age[infos$isAD & infos$included]),
                                 mean(infos$age[infos$isHC & infos$included]))),
                     age_sd=round(c(sd(infos$age[infos$isSD & infos$included]),
                                    sd(infos$age[infos$isAD & infos$included]),
                                    sd(infos$age[infos$isHC & infos$included])))
)

print(d.demog)
```

## DVARS analysis
```{r DVARS analysis, fig.height=4, fig.width=4}
a = read.table(file.path(PATH, 'DVARS_vols_affected.txt'), header = F)
b = cbind(rep(0,N), a[,1:ncol(a)-1])
affected = a | b
rm(a); rm(b)

# which subjetc have more than 20% of their volumes affected?
print(as.character(infos$id[rowSums(affected) > infos$N_vols*0.2]))

lillie.test(infos$SD_SDVARS_X2)
kruskal.test(SD_SDVARS_X2 ~ group, data=infos)

lillie.test(infos$g_Avar)
kruskal.test(g_Avar ~ group, data=infos)

# DVARS after excluding subjects
infos_ = subset(infos, included)
kruskal.test(SD_SDVARS_X2 ~ group, data=infos_)
kruskal.test(g_Avar ~ group, data=infos_)

p1 = ggplot(infos, aes(x=group, y=SD_SDVARS_X2)) + geom_violin() + geom_boxplot(width=0.4) + ylim(c(0.5,2)) +
  geom_point(shape=1, color="gray70", size=4, position = position_jitter(width = 0.2, height = 0.0002)) +
  ggtitle("before exclusion")

p2 = ggplot(infos_, aes(x=group, y=SD_SDVARS_X2)) + geom_violin() + geom_boxplot(width=0.4) + ylim(c(0.5,2)) +
  geom_point(shape=1, color="gray70", size=4, position = position_jitter(width = 0.2, height = 0.0002)) + 
  ggtitle("after exclusion")

p3 = ggplot(infos, aes(x=group, y=g_Avar)) + geom_violin() + geom_boxplot(width=0.4) +
  geom_point(shape=1, color="gray70", size=4, position = position_jitter(width = 0.2, height = 0.0002)) +
  ggtitle("before exclusion")

p4 = ggplot(infos_, aes(x=group, y=g_Avar)) + geom_violin() + geom_boxplot(width=0.4) +
  geom_point(shape=1, color="gray70", size=4, position = position_jitter(width = 0.2, height = 0.0002)) + 
  ggtitle("after exclusion")

plot_grid(p1, p2, p3, p4, ncol = 2, nrow = 2, rel_widths = c(1, 1))
```


## Load time series of 200 ROIs, global signal correction
```{r Load time series of 200 ROIs GSC, warning=FALSE}
# first stockholm sample, then china sample
f1 = list.files(path = PATH_ROIS200, pattern = "^ADSD.*.txt")
f2 = list.files(path = PATH_ROIS200, pattern = "^20.*.txt")
f = c(f1,f2)
rm(f1); rm(f2)
assert(length(f) == N)

R = array(NA, dim=c(N_ROIS[2], N_ROIS[2], N))
P = array(NA, dim=c(N_ROIS[2], N_ROIS[2], N))
R_gs = array(NA, dim=c(N_ROIS[2], N_ROIS[2], N))
P_gs = array(NA, dim=c(N_ROIS[2], N_ROIS[2], N))
for (s in 1:N) {
  d = read.table(file.path(PATH_ROIS200, f[s]), header = F)
  idx = !affected[s,][1:infos$N_vols[s]]
  d = as.matrix(d[idx,]) # excluding affected scans
  tmp =  rcorr(d, type="pearson")
  R[,,s] = tmp$r
  P[,,s] = tmp$P
  
  # global signal regression
  d_gs = array(NA, dim=dim(d))
  gs = apply(d, 1, mean, na.rm=T)
  for (r in 1:N_ROIS[2]) {
    if (sum(is.na(d[,r])) == length(gs)) {
      d_gs[,r] = rep(NA, length(gs))
    } else {
      fit = lm(d[,r] ~ gs)
      d_gs[,r] = fit$residuals
    }
  }
  
  tmp =  rcorr(d_gs, type="pearson")
  R_gs[,,s] = tmp$r
  P_gs[,,s] = tmp$P
}

```


## Mean correlation per group 200 ROIs
```{r Mean correlation per group 200 ROIs}
# Correlation R
reidx=1:N_ROIS[2]
tri = lower.tri(array(0, dim=c(N_ROIS[2], N_ROIS[2])))

FC200 = list()
FC200$R$SD = apply(R[,,infos$isSD & infos$included], c(1,2), mean)
FC200$R$AD = apply(R[,,infos$isAD & infos$included], c(1,2), mean, na.rm=T) # problem with ADSD_13
FC200$R$HC = apply(R[,,infos$isHC & infos$included], c(1,2), mean)

# P values
FC200$P$SD = apply(P[,,infos$isSD & infos$included], c(1,2), mean)
FC200$P$AD = apply(P[,,infos$isAD & infos$included], c(1,2), mean, na.rm=T) # problem with ADSD_13
FC200$P$HC = apply(P[,,infos$isHC & infos$included], c(1,2), mean)

# FDR 0.05
tmp = rbind(FC200$P$SD, FC200$P$AD, FC200$P$HC)
tmp[!rbind(tri,tri,tri)]=NA

# Global FDR across groups
M = array(p.adjust(tmp, method = "fdr"), dim=c(3*N_ROIS[2], N_ROIS[2]))

FC200$R_FDR05$SD = FC200$R$SD
FC200$R_FDR05$AD = FC200$R$AD
FC200$R_FDR05$HC = FC200$R$HC

FC200$P_FDR05$SD = M[1:200,]
FC200$P_FDR05$AD = M[201:400,]
FC200$P_FDR05$HC = M[401:600,]

FC200$R_FDR05$SD[FC200$P_FDR05$SD >= 0.05 | is.na(FC200$P_FDR05$AD)] = 0
FC200$R_FDR05$AD[FC200$P_FDR05$AD >= 0.05 | is.na(FC200$P_FDR05$AD)] = 0
FC200$R_FDR05$HC[FC200$P_FDR05$HC >= 0.05 | is.na(FC200$P_FDR05$HC)] = 0
```

## Plot: Pearson's r 200 ROIs
```{r Plot: Pearson r 200 ROIs, fig.height=15, fig.width=12}
limR = c(-1, 1)
p1 = ggplot(melt(rmna(FC200$R$SD[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("SD")

p2 = ggplot(melt(rmna(FC200$R_FDR05$SD[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("FDR 0.05")

p3 = ggplot(melt(rmna(FC200$R$AD[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("AD")

p4 = ggplot(melt(rmna(FC200$R_FDR05$AD[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("FDR 0.05")

p5 = ggplot(melt(rmna(FC200$R$HC[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("FDR 0.05")

p6 = ggplot(melt(rmna(FC200$R_FDR05$HC[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("FDR 0.05")

plot_grid(p1, p2, p3, p4, p5, p6, ncol = 2, nrow = 3, rel_widths = c(1, 1))
```

## Fisher z-transformation 200 ROIs
```{r Fisher z-transformation 200 ROIs}
Z = 0.5*(log(1+R) - log(1-R))
```

## Load ROIs of temporal regions
```{r Load ROIs}
tmpidx=as.matrix(read.table(file.path(PATH, 'Cradd_tmp_labels.txt'), header = F))
tmpidx=tmpidx[1,]
N_ROIS[3]=length(tmpidx)
```


## t-test
```{r t-test}
HCvsSD=list()
HCvsAD=list()

HCvsSD$Tvals = array(NA, dim=c(N_ROIS[3], N_ROIS[3]))
HCvsSD$Pvals = array(NA, dim=c(N_ROIS[3], N_ROIS[3]))

HCvsAD$Tvals = array(NA, dim=c(N_ROIS[3], N_ROIS[3]))
HCvsAD$Pvals = array(NA, dim=c(N_ROIS[3], N_ROIS[3]))

for (i in 1:N_ROIS[3]) {
  for (j in 1:N_ROIS[3]) {
    if (i > j ) {
      
      y = Z[tmpidx[i],tmpidx[j],]
      
      # SD
      idx = (infos$isSD | infos$isHC) & infos$included
      result = t.test(y[idx] ~ infos$group[idx], var.equal=T,
                      alternative = "greater")
      HCvsSD$Tvals[i,j] = result[[1]]
      HCvsSD$Pvals[i,j] = result[[3]]
      
      # AD
      idx = (infos$isAD | infos$isHC) & infos$included
      result = t.test(y[idx] ~ infos$group[idx], var.equal=T,
                      alternative = "less")
      HCvsAD$Tvals[i,j] = -result[[1]]
      HCvsAD$Pvals[i,j] = result[[3]]
      
    }
  }
}

# FDR correction
tmp = rbind(HCvsSD$Pvals, HCvsAD$Pvals)
M = array(p.adjust(tmp, method = "fdr"), dim=c(2*N_ROIS[3], N_ROIS[3]))

HCvsSD$Pvals_FDR05 = M[1:N_ROIS[3],]
HCvsAD$Pvals_FDR05 = M[(N_ROIS[3]+1):(N_ROIS[3]*2),]

HCvsSD$Tvals_FDR05 = HCvsSD$Tvals
HCvsAD$Tvals_FDR05 = HCvsAD$Tvals

# sum(HCvsSD$Pvals_FDR05 < 0.05, na.rm = T)
# sum(HCvsAD$Pvals_FDR05 < 0.05, na.rm = T)

HCvsSD$Tvals_FDR05[HCvsSD$Pvals_FDR05 >= 0.05 | is.na(HCvsSD$Pvals_FDR05)] = 0
HCvsAD$Tvals_FDR05[HCvsAD$Pvals_FDR05 >= 0.05 | is.na(HCvsAD$Pvals_FDR05)] = 0

# max(rbind(HCvsSD$Tvals,HCvsAD$Tvals), na.rm=T)

```

## Plot: t-test
```{r Plot: t-test, fig.height=12, fig.width=15}
reidx=1:N_ROIS[3]
tri = lower.tri(array(0, dim=c(N_ROIS[3], N_ROIS[3])))

mylim = c(-5.3, 5.3)

p1 = ggplot(melt(rmna(HCvsSD$Tvals[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = mylim, space = "Lab", name = "t") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > SD")

p2 = ggplot(melt(HCvsSD$Tvals_FDR05[reidx,reidx]), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = mylim, space = "Lab", name = "t") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > SD (FDR 0.05)")

p3 = ggplot(melt(rmna(HCvsAD$Tvals[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = mylim, space = "Lab", name = "t") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > AD")

p4 = ggplot(melt(HCvsAD$Tvals_FDR05[reidx,reidx]), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = mylim, space = "Lab", name = "t") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > AD (FDR 0.05)")

plot_grid(p1,p2,p3,p4, ncol = 2, nrow = 2, rel_widths = c(1, 1))
```


## Mean correlation per group 200 ROIs after global sign. reg.
```{r Mean correlation per group 200 ROIs global sig reg}
# Correlation R
FC200gs = list()
FC200gs$R$SD = apply(R_gs[,,infos$isSD & infos$included], c(1,2), mean)
FC200gs$R$AD = apply(R_gs[,,infos$isAD & infos$included], c(1,2), mean, na.rm=T) # problem with ADSD_13
FC200gs$R$HC = apply(R_gs[,,infos$isHC & infos$included], c(1,2), mean)

# P values
FC200gs$P$SD = apply(P_gs[,,infos$isSD & infos$included], c(1,2), mean)
FC200gs$P$AD = apply(P_gs[,,infos$isAD & infos$included], c(1,2), mean, na.rm=T) # problem with ADSD_13
FC200gs$P$HC = apply(P_gs[,,infos$isHC & infos$included], c(1,2), mean)

# FDR 0.05
tmp = rbind(FC200gs$P$SD, FC200gs$P$AD, FC200gs$P$HC)
tmp[!rbind(tri,tri,tri)]=NA

# Global FDR across groups
M = array(p.adjust(tmp, method = "fdr"), dim=c(3*N_ROIS[2], N_ROIS[2]))

FC200gs$R_FDR05$SD = FC200gs$R$SD
FC200gs$R_FDR05$AD = FC200gs$R$AD
FC200gs$R_FDR05$HC = FC200gs$R$HC

FC200gs$P_FDR05$SD = M[1:200,]
FC200gs$P_FDR05$AD = M[201:400,]
FC200gs$P_FDR05$HC = M[401:600,]

FC200gs$R_FDR05$SD[FC200gs$P_FDR05$SD >= 0.05 | is.na(FC200gs$P_FDR05$SD)] = 0
FC200gs$R_FDR05$AD[FC200gs$P_FDR05$AD >= 0.05 | is.na(FC200gs$P_FDR05$AD)] = 0
FC200gs$R_FDR05$HC[FC200gs$P_FDR05$HC >= 0.05 | is.na(FC200gs$P_FDR05$HC)] = 0

reidx=1:N_ROIS[2]
tri = lower.tri(array(0, dim=c(N_ROIS[2], N_ROIS[2])))
```

## Plot: FC 200 ROIs after global sign. reg.
```{r Plot: FC 200 ROIs global sig reg, fig.height=15, fig.width=6}
limR = c(-1, 1)
p1 = ggplot(melt(rmna(FC200gs$R$SD[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("SD")

p2 = ggplot(melt(rmna(FC200gs$R_FDR05$SD[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("FDR 0.05")

p3 = ggplot(melt(rmna(FC200gs$R$AD[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("AD")

p4 = ggplot(melt(rmna(FC200gs$R_FDR05$AD[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("FDR 0.05")

p5 = ggplot(melt(rmna(FC200gs$R$HC[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("FDR 0.05")

p6 = ggplot(melt(rmna(FC200gs$R_FDR05$HC[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("FDR 0.05")

plot_grid(p1, p3, p5, ncol = 1, nrow = 3)
```

## Count sign. edges
```{r Count sign edges}
l=(c("SD", "AD", "HC"))
x1 = c(sum(FC200$R_FDR05$SD*tri > 0, na.rm = T), sum(FC200$R_FDR05$AD*tri > 0),
       sum(FC200$R_FDR05$HC*tri > 0))
sprintf("%s %d %d%%", l, x1, round(x1/x1[3]*100))
x2 = c(sum(FC200gs$R_FDR05$SD*tri > 0, na.rm = T), sum(FC200gs$R_FDR05$AD*tri > 0),
       sum(FC200gs$R_FDR05$HC*tri > 0))
sprintf("%s %d %d%%", l, x2, round(x2/x2[3]*100))

```

## Fisher z-transformation 200 ROIS after global sign. reg.
```{r Fisher z-transformation 200 ROIS global sign reg}
Z = 0.5*(log(1+R_gs) - log(1-R_gs))
```

## t-test after global sign. reg.
```{r t-test after global sign reg}
# HCvsSD=list()
# HCvsAD=list()
# 
# HCvsSD$Tvals = array(NA, dim=c(N_ROIS[3], N_ROIS[3]))
# HCvsSD$Pvals = array(NA, dim=c(N_ROIS[3], N_ROIS[3]))
# HCvsAD$Tvals = array(NA, dim=c(N_ROIS[3], N_ROIS[3]))
# HCvsAD$Pvals = array(NA, dim=c(N_ROIS[3], N_ROIS[3]))
# 
# HCvsSD$perm.Svals = array(NA, dim=c(N_ROIS[3], N_ROIS[3]))
# HCvsSD$perm.Pvals = array(NA, dim=c(N_ROIS[3], N_ROIS[3]))
# HCvsAD$perm.Svals = array(NA, dim=c(N_ROIS[3], N_ROIS[3]))
# HCvsAD$perm.Pvals = array(NA, dim=c(N_ROIS[3], N_ROIS[3]))
# 
# for (i in 1:N_ROIS[3]) {
#   print(i)
#   for (j in 1:N_ROIS[3]) {
#     if (i > j) {
# 
#       y = Z[tmpidx[i], tmpidx[j], ]
# 
#       # SD
#       idx = (infos$isSD | infos$isHC) & infos$included
# 
#       result = t.test(y[idx] ~ infos$group[idx],
#                       var.equal = T,
#                       alternative = "greater")
# 
#       HCvsSD$Tvals[i, j] = result[[1]]
#       HCvsSD$Pvals[i, j] = result[[3]]
# 
#       resp = permTS(y[idx] ~ infos$group[idx], alternative = "greater", method = "exact.mc",
#                     control = permControl(nmc = 5000, setSEED = FALSE))
# 
#       HCvsSD$perm.Svals[i, j] = resp$estimate
#       HCvsSD$perm.Pvals[i, j] = resp$p.value
# 
#       # AD
#       idx = (infos$isAD | infos$isHC) & infos$included
# 
#       result = t.test(y[idx] ~ infos$group[idx],
#                       var.equal = T,
#                       alternative = "less")
# 
#       HCvsAD$Tvals[i, j] = -result[[1]]
#       HCvsAD$Pvals[i, j] =  result[[3]] #resp$p.value
# 
#       resp = permTS(y[idx] ~ infos$group[idx], alternative = "less", method = "exact.mc",
#                     control = permControl(nmc = 5000, setSEED = FALSE))
# 
#       HCvsAD$perm.Svals[i, j] = -resp$estimate
#       HCvsAD$perm.Pvals[i, j] = resp$p.value
#     }
#   }
# }
# 
# save(HCvsSD, HCvsAD, file = file.path(PATH, 'ttest.RData'))
```

## Load t-test and permutation test
```{r load t-test and permutation test}
load(file.path(PATH, 'ttest.RData'))
```

## FDR Correction
```{r FDR correction}
# FDR correction
tmp = rbind(HCvsSD$Pvals, HCvsAD$Pvals)
M = array(p.adjust(tmp, method = "fdr"), dim=c(2*N_ROIS[3], N_ROIS[3]))

HCvsSD$Pvals_FDR05 = M[1:N_ROIS[3],]
HCvsAD$Pvals_FDR05 = M[(N_ROIS[3]+1):(N_ROIS[3]*2),]

HCvsSD$Tvals_FDR05 = HCvsSD$Tvals
HCvsAD$Tvals_FDR05 = HCvsAD$Tvals

HCvsSD$Tvals_FDR05[HCvsSD$Pvals_FDR05 >= 0.05 | is.na(HCvsSD$Pvals_FDR05)] = 0
HCvsAD$Tvals_FDR05[HCvsAD$Pvals_FDR05 >= 0.05 | is.na(HCvsAD$Pvals_FDR05)] = 0

# permutations
tmp = rbind(HCvsSD$perm.Pvals, HCvsAD$perm.Pvals)
M = array(p.adjust(tmp, method = "fdr"), dim=c(2*N_ROIS[3], N_ROIS[3]))

HCvsSD$perm.Pvals_FDR05 = M[1:N_ROIS[3],]
HCvsAD$perm.Pvals_FDR05 = M[(N_ROIS[3]+1):(N_ROIS[3]*2),]

HCvsSD$perm.Svals_FDR05 = HCvsSD$perm.Svals
HCvsAD$perm.Svals_FDR05 = HCvsAD$perm.Svals

HCvsSD$perm.Svals_FDR05[HCvsSD$perm.Pvals_FDR05 >= 0.062 | is.na(HCvsSD$perm.Pvals_FDR05)] = 0
HCvsAD$perm.Svals_FDR05[HCvsAD$perm.Pvals_FDR05 >= 0.062 | is.na(HCvsAD$perm.Pvals_FDR05)] = 0

```

## Plot: t-Test after global sign. reg.
```{r Plot: t-Test after global sign reg, fig.height=12, fig.width=15}
reidx=1:N_ROIS[3]
tri = lower.tri(array(0, dim=c(N_ROIS[3], N_ROIS[3])))

mylim = c(-5, 5)

p1 = ggplot(melt(rmna(HCvsSD$Tvals[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = mylim, space = "Lab", name = "t") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > SD")

p2 = ggplot(melt(HCvsSD$Tvals_FDR05[reidx,reidx]), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = mylim, space = "Lab", name = "t") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > SD (FDR 0.05)")

p3 = ggplot(melt(rmna(HCvsAD$Tvals[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = mylim, space = "Lab", name = "t") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > AD")

p4 = ggplot(melt(HCvsAD$Tvals_FDR05[reidx,reidx]), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = mylim, space = "Lab", name = "t") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > AD (FDR 0.05)")

plot_grid(p1,p2,p3,p4, ncol = 2, nrow = 2, rel_widths = c(1, 1))
```

## Plot: permutation test after global sign. reg.
```{r Plot: permutation test after global sign reg, fig.height=12, fig.width=15}
reidx=1:N_ROIS[3]
tri = lower.tri(array(0, dim=c(N_ROIS[3], N_ROIS[3])))

mylim = c(-0.5, 0.5)

p1 = ggplot(melt(rmna(HCvsSD$perm.Svals[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = mylim, space = "Lab", name = "z diff") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > SD")

p2 = ggplot(melt(HCvsSD$perm.Svals_FDR05[reidx,reidx]), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = mylim, space = "Lab", name = "z diff") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > SD (FDR 0.05)")

p3 = ggplot(melt(rmna(HCvsAD$perm.Svals[reidx,reidx])*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = mylim, space = "Lab", name = "z diff") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > AD")

p4 = ggplot(melt(HCvsAD$perm.Svals_FDR05[reidx,reidx]), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = mylim, space = "Lab", name = "z diff") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > AD (FDR 0.05)")

plot_grid(p1,p2,p3,p4, ncol = 2, nrow = 2, rel_widths = c(1, 1))
```

## Significant edges
### HC > SD
```{r Significant edges HC vs SD}
rois = read.table(file.path(PATH, 'ROIs_22.txt'), header = T)
map  = read.table(file.path(PATH, 'Cradd_AAL_mapping.txt'), header = T)

idx = which(HCvsSD$Pvals_FDR05 < 0.05, arr.ind = T)
tvalues = HCvsSD$Tvals_FDR05[idx]
orderSD = order(tvalues, decreasing = T)
edgesSD = array(tmpidx[idx], dim=c(nrow(idx),2))
print(edgesSD[orderSD,])
write(edgesSD, file=file.path(PATH, 'SDedges.txt'), ncolumns = 1)

# map craddock numbers to AAL labels
resultSD = array(character(nrow(edgesSD)*2), c(nrow(edgesSD),2))
for (i in 1:nrow(edgesSD)) {
  resultSD[i,] = c(as.character(rois$AAL_Name[map$aal[match(edgesSD[i,1], map$crad)]]),
                   as.character(rois$AAL_Name[map$aal[match(edgesSD[i,2], map$crad)]]))
}
print(cbind(resultSD[orderSD,], sprintf('%.2f',tvalues[orderSD])))
```

### HC > AD
```{r Significant edges HC vs AD}
idx = which(HCvsAD$Pvals_FDR05 < 0.05, arr.ind = T)
edgesAD = array(tmpidx[idx], dim=c(nrow(idx),2))
tvalues = HCvsAD$Tvals_FDR05[idx]
orderAD = order(tvalues, decreasing = T)
print(edgesAD[orderAD,])
write(edgesAD, file=file.path(PATH, 'ADedges.txt'), ncolumns = 1)

# map craddock numbers to AAL labels
resultAD = array(character(nrow(edgesAD)*2), c(nrow(edgesAD),2))
for (i in 1:nrow(edgesAD)) {
  resultAD[i,] = c(as.character(rois$AAL_Name[map$aal[match(edgesAD[i,1], map$crad)]]),
                 as.character(rois$AAL_Name[map$aal[match(edgesAD[i,2], map$crad)]]))
}
print(cbind(resultAD[orderAD,], sprintf('%.2f',tvalues[orderAD])))
```

## Z-scores SD
```{r Z-scores SD, warning=FALSE, fig.height=4, fig.width=7}
zSD = array(NA, dim=c(N_SD, nrow(edgesSD)))
zHC = array(NA, dim=c(N_HC, nrow(edgesSD)))

p=list()
for (i in 1:nrow(edgesSD)) {
  zSD[,i] = Z[edgesSD[i,1], edgesSD[i,2], infos$isSD & infos$included]
  zHC[,i] = Z[edgesSD[i,1], edgesSD[i,2], infos$isHC & infos$included]
  
  d=data.frame(z=c(zSD[,i], zHC[,i]), group=as.factor(c(rep("SD", N_SD), rep("HC", N_HC))),
               med = c(rep(median(zSD[,i]), N_SD), rep(median(zHC[,i]), N_HC)))
  
  p[[i]]= ggplot(d, aes(x=group, y=z)) + geom_violin() + geom_boxplot(width=0.4) +
    geom_point(shape=1, color="gray60", size=1.5, position = position_jitter(width = 0.2, height = 0.0002)) + 
    geom_line(aes(x=as.numeric(group), y=med), size=0.3) + coord_cartesian(ylim = c(-0.5, 1)) +
    ggtitle(sprintf('%s %d\n%s %d', resultSD[i,1], edgesSD[i,1], resultSD[i,2], edgesSD[i,2])) +
    theme(plot.title = element_text(size=10, face = "plain"))
}

p = p[orderSD]

plot_grid(p[[1]], p[[2]], p[[3]], p[[4]], p[[5]], p[[6]], p[[7]], p[[8]], p[[9]], p[[10]], 
          ncol = 5, nrow = 2)

```

## Z-scores AD
```{r Z-scores AD, warning=FALSE, fig.height=2, fig.width=3}
zAD = array(NA, dim=c(sum(infos$isAD & infos$included), nrow(edgesAD)))
zHC = array(NA, dim=c(sum(infos$isHC & infos$included), nrow(edgesAD)))

p=list()
for (i in 1:nrow(edgesAD)) {
  zAD[,i] = Z[edgesAD[i,1], edgesAD[i,2], infos$isAD & infos$included]
  zHC[,i] = Z[edgesAD[i,1], edgesAD[i,2], infos$isHC & infos$included]
  
  d=data.frame(z=c(zAD[,i], zHC[,i]), group=as.factor(c(rep("AD", N_AD), rep("HC", N_HC))),
               med = c(rep(median(zAD[,i]), N_AD), rep(median(zHC[,i]), N_HC)))
  
  p[[i]]= ggplot(d, aes(x=group, y=z)) + geom_violin() + geom_boxplot(width=0.4) +
    geom_point(shape=1, color="gray60", size=1.5, position = position_jitter(width = 0.2, height = 0.0002)) + 
    geom_line(aes(x=as.numeric(group), y=med), size=0.3) + coord_cartesian(ylim = c(-0.5, 1)) +
    ggtitle(sprintf('%s %d\n%s %d', resultAD[i,1], edgesAD[i,1], resultAD[i,2], edgesAD[i,2])) +
    theme(plot.title = element_text(size=10, face = "plain"))
}

p = p[orderAD]

plot_grid(p[[1]], p[[2]], ncol = 2, nrow = 1)
```

## Significant edges
### HC > SD
```{r Significant edges HC vs SD perm}
rois = read.table(file.path(PATH, 'ROIs_22.txt'), header = T)
map  = read.table(file.path(PATH, 'Cradd_AAL_mapping.txt'), header = T)

idx = which(HCvsSD$perm.Pvals_FDR05 < 0.062, arr.ind = T)
tvalues = HCvsSD$perm.Svals_FDR05[idx]
orderSD = order(tvalues, decreasing = T)
edgesSD = array(tmpidx[idx], dim=c(nrow(idx),2))
print(edgesSD[orderSD,])
write(edgesSD, file=file.path(PATH, 'SDedges_perm.txt'), ncolumns = 1)

# map craddock numbers to AAL labels
resultSD = array(character(nrow(edgesSD)*2), c(nrow(edgesSD),2))
for (i in 1:nrow(edgesSD)) {
  resultSD[i,] = c(as.character(rois$AAL_Name[map$aal[match(edgesSD[i,1], map$crad)]]),
                   as.character(rois$AAL_Name[map$aal[match(edgesSD[i,2], map$crad)]]))
}
print(cbind(resultSD[orderSD,], sprintf('%.2f',tvalues[orderSD])))
```

### HC > AD
```{r Significant edges HC vs AD perm}
idx = which(HCvsAD$perm.Pvals_FDR05 < 0.062, arr.ind = T)
edgesAD = array(tmpidx[idx], dim=c(nrow(idx),2))
tvalues = HCvsAD$perm.Svals_FDR05[idx]
orderAD = order(tvalues, decreasing = T)
print(edgesAD[orderAD,])
write(edgesAD, file=file.path(PATH, 'ADedges_perm.txt'), ncolumns = 1)

# map craddock numbers to AAL labels
resultAD = array(character(nrow(edgesAD)*2), c(nrow(edgesAD),2))
for (i in 1:nrow(edgesAD)) {
  resultAD[i,] = c(as.character(rois$AAL_Name[map$aal[match(edgesAD[i,1], map$crad)]]),
                 as.character(rois$AAL_Name[map$aal[match(edgesAD[i,2], map$crad)]]))
}
print(cbind(resultAD[orderAD,], sprintf('%.2f',tvalues[orderAD])))
```

## Z-scores SD
```{r Z-scores SD perm, warning=FALSE, fig.height=4, fig.width=8}
zSD = array(NA, dim=c(sum(infos$isSD & infos$included), nrow(edgesSD)))
zHC = array(NA, dim=c(sum(infos$isHC & infos$included), nrow(edgesSD)))

p=list()
for (i in 1:nrow(edgesSD)) {
  zSD[,i] = Z[edgesSD[i,1], edgesSD[i,2], infos$isSD & infos$included]
  zHC[,i] = Z[edgesSD[i,1], edgesSD[i,2], infos$isHC & infos$included]
  
  d=data.frame(z=c(zSD[,i], zHC[,i]), group=as.factor(c(rep("SD", N_SD), rep("HC", N_HC))),
                med = c(rep(median(zSD[,i]), N_SD), rep(median(zHC[,i]), N_HC)))
  
  p[[i]]= ggplot(d, aes(x=group, y=z)) + geom_violin() + geom_boxplot(width=0.4) +
    geom_point(shape=1, color="gray60", size=1.5, position = position_jitter(width = 0.2, height = 0.0002)) +
    geom_line(aes(x=as.numeric(group), y=med), size=0.3) + coord_cartesian(ylim = c(-0.5, 1)) +
    ggtitle(sprintf('%s %d\n%s %d', resultSD[i,1], edgesSD[i,1], resultSD[i,2], edgesSD[i,2])) +
    theme(plot.title = element_text(size=10, face = "plain"))
}

p = p[orderSD]

plot_grid(p[[1]], p[[2]], p[[3]], p[[4]], p[[5]], p[[6]], p[[7]], p[[8]], p[[9]], p[[10]], p[[11]],
          p[[12]], ncol = 6, nrow = 2)

```

## Z-scores AD
```{r Z-scores AD perm, warning=FALSE, fig.height=2, fig.width=3}
zAD = array(NA, dim=c(sum(infos$isAD & infos$included), nrow(edgesAD)))
zHC = array(NA, dim=c(sum(infos$isHC & infos$included), nrow(edgesAD)))

p=list()
for (i in 1:nrow(edgesAD)) {
  zAD[,i] = Z[edgesAD[i,1], edgesAD[i,2], infos$isAD & infos$included]
  zHC[,i] = Z[edgesAD[i,1], edgesAD[i,2], infos$isHC & infos$included]
  
  d=data.frame(z=c(zAD[,i], zHC[,i]), group=as.factor(c(rep("AD", N_AD), rep("HC",N_HC))),
               med = c(rep(median(zAD[,i]), N_AD), rep(median(zHC[,i]), N_HC)))
  
  p[[i]]= ggplot(d, aes(x=group, y=z)) + geom_violin() + geom_boxplot(width=0.4) +
    geom_point(shape=1, color="gray60", size=1.5, position = position_jitter(width = 0.2, height = 0.0002)) + 
    geom_line(aes(x=as.numeric(group), y=med), size=0.3) + coord_cartesian(ylim = c(-0.5, 1)) +
    ggtitle(sprintf('%s %d\n%s %d', resultAD[i,1], edgesAD[i,1], resultAD[i,2], edgesAD[i,2])) +
    theme(plot.title = element_text(size=10, face = "plain"))
}

p = p[orderAD]

plot_grid(p[[1]], p[[2]], ncol = 2, nrow = 1)
```



