---
title: "ADSD"
author: "Simon Schwab"
date: "10 May 2017"
output: html_document
---

## Install R Markdown 
```{r Install packages}
# install.packages("rmarkdown")
# install.packages("testit")
# install.packages("Hmisc")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r Load libraries, message=FALSE, warning=TRUE}
library(testit)
library(reshape2)
library(ggplot2)
library(cowplot)
library(Hmisc)
```

## Main Variables
```{r Main variables}
N = 62
N_ROIS = 14
PATH_HOME = "/home/simon"
PATH_ROIS = file.path(PATH_HOME, 'Drive', '14RoisN62') 
PATH = file.path(PATH_HOME, 'Data', 'ADSD')
```

## Subject infos
```{r Subjects infos}
infos = read.table(file.path(PATH, 'infos.txt'), header = T)
rois = read.table(file.path(PATH, 'ROIs.txt'), header = T)
# Excluded subjects
infos$included = rep(TRUE,N)
# DVARS; more than 20% volumes affected:
infos$included[match("ADSD_13", infos$id)] = FALSE
infos$included[match("ADSD_36", infos$id)] = FALSE
# High E-vars:
infos$included[match("ADSD_34", infos$id)] = FALSE
# DVARS below 40%:
infos$included[match("ADSD_15", infos$id)] = FALSE
infos$included[match("ADSD_42", infos$id)] = FALSE
infos$included[match("ADSD_43", infos$id)] = FALSE

infos$included[match("2026", infos$id)] = FALSE # no mri data

infos$isHC = as.logical(infos$isHC)
infos$isAD = as.logical(infos$isAD)
infos$isSD = as.logical(infos$isSD)

infos$group=rep(NA,N)
infos$group[infos$isSD] = 'SD'
infos$group[infos$isAD] = 'AD'
infos$group[infos$isHC] = 'HC'
infos$group = as.factor(infos$group)

assert(nrow(infos) == N)
N_SD  = sum(infos$isSD)
N_AD  = sum(infos$isAD)
N_HC  = sum(infos$isHC)
assert(N_SD + N_AD + N_HC == N)

print(infos)
```
## Demographics
```{r Demographics}
d.demog = data.frame(Group=as.factor(c('SD', 'AD', 'HC')),
                     N_Total=c(sum(infos$isSD), sum(infos$isAD), sum(infos$isHC)),
                     N_included=c(sum(infos$isSD & infos$included), sum(infos$isAD & infos$included),
                                  sum(infos$isHC & infos$included)),
                     age_mean=round(c(mean(infos$age[infos$isSD & infos$included]),
                                 mean(infos$age[infos$isAD & infos$included]),
                                 mean(infos$age[infos$isHC & infos$included]))),
                     age_sd=round(c(sd(infos$age[infos$isSD & infos$included]),
                                    sd(infos$age[infos$isAD & infos$included]),
                                    sd(infos$age[infos$isHC & infos$included])))
)

print(d.demog)
```

## ROI labels
Time series are extracted using the k=200 Craddock atlas with a selection of ROIS most prominently overlapping with tempral AAL regions. `AAL_Nr` indicates atlas label, `position` indicates position in matrices from figures and plots. In these, first uneven (left), then even (right) regions are shown.

```{r ROI labels}
rois$position = c(t(array(c(1:N_ROIS), dim=c(7,2))))
print(rois)

# reorder ROIS, first left (uneven) then right (even) rois
reidx=c(seq(1,N_ROIS,2), seq(2,N_ROIS,2))

# lower diagonal
tri = lower.tri(array(0, dim=c(N_ROIS, N_ROIS)))
```

## DVARS analysis
```{r DVARS analysis}
a = read.table(file.path(PATH, 'DVARS_vols_affected.txt'), header = F)
b = cbind(rep(0,N), a[,1:ncol(a)-1])
affected = a | b

# which subjetc have more than 20% of their volumes affected?
print(as.character(infos$id[rowSums(affected) > infos$N_vols*0.2]))
```
DVARS diagnostics

```{r DVARS diagnostics}
dvars=read.table(file.path(PATH, 'DVARS.txt'), header = T)
```

## Load time series of 14 Rois
```{r Load time series of 14 Rois, warning=FALSE}
# first stockholm sample, then china sample
f1 = list.files(path = PATH_ROIS, pattern = "^ADSD.*.txt")
f2 = list.files(path = PATH_ROIS, pattern = "^20.*.txt")
f = c(f1,f2)
rm(f1); rm(f2)
assert(length(f) == N)

R = array(NA, dim=c(N_ROIS, N_ROIS, N))
P = array(NA, dim=c(N_ROIS, N_ROIS, N))
for (s in 1:N) {
  d = read.table(file.path(PATH_ROIS, f[s]), header = F)
  # R[,,s] = cor(d)
  idx = !affected[s,][1:infos$N_vols[s]] # excluding affected scans
  tmp =  rcorr(as.matrix(d[idx,]), type="pearson")
  
  R[,,s] = tmp$r
  P[,,s] = tmp$P
}
# ADSD_13 has no time series from ROI 11 and ROI 12
```


## Mean correlation per group
```{r Mean correlation per group}
# Correlation R
meanRSD = apply(R[,,infos$isSD & infos$included], c(1,2), mean)
meanRAD = apply(R[,,infos$isAD & infos$included], c(1,2), mean, na.rm=T) # problem with ADSD_13
meanRHC = apply(R[,,infos$isHC & infos$included], c(1,2), mean)

# P values
meanPSD = apply(P[,,infos$isSD & infos$included], c(1,2), mean)
meanPAD = apply(P[,,infos$isAD & infos$included], c(1,2), mean, na.rm=T) # problem with ADSD_13
meanPHC = apply(P[,,infos$isHC & infos$included], c(1,2), mean)

# p = 0.01 uncorrected
meanRSD_01 = meanRSD
meanRSD_01[meanPSD > 0.01] = 0
meanRAD_01 = meanRAD
meanRAD_01[meanPAD > 0.01] = 0
meanRHC_01 = meanRHC
meanRHC_01[meanPHC > 0.01] = 0

# FDR 0.05
meanPSD_ = meanPSD[reidx,reidx]
meanPAD_ = meanPAD[reidx,reidx]
meanPHC_ = meanPHC[reidx,reidx]

meanRSD_FDR05 = meanRSD[reidx,reidx]
meanRAD_FDR05 = meanRAD[reidx,reidx]
meanRHC_FDR05 = meanRHC[reidx,reidx]

meanPSD_[!tri] = NA
meanPAD_[!tri] = NA
meanPHC_[!tri] = NA

#meanPSD_FDR05 = array(p.adjust(meanPSD_, method = "fdr"), dim=c(N_ROIS,N_ROIS))
#meanPAD_FDR05 = array(p.adjust(meanPAD_, method = "fdr"), dim=c(N_ROIS,N_ROIS))
#meanPHC_FDR05 = array(p.adjust(meanPHC_, method = "fdr"), dim=c(N_ROIS,N_ROIS))

# Global FDR across groups
M = array(p.adjust(rbind(meanPSD_, meanPAD_, meanPHC_), method = "fdr"), 
          dim=c(3*N_ROIS, N_ROIS))
meanPSD_FDR05 = M[1:14,]
meanPAD_FDR05 = M[15:28,]
meanPHC_FDR05 = M[29:42,]

meanRSD_FDR05[meanPSD_FDR05 > 0.05 | is.na(meanPSD_FDR05)] = 0
meanRAD_FDR05[meanPAD_FDR05 > 0.05 | is.na(meanPAD_FDR05)] = 0
meanRHC_FDR05[meanPHC_FDR05 > 0.05 | is.na(meanPHC_FDR05)] = 0
```

## Plot: FC
```{r Plot: FC, fig.height=4.5, fig.width=9}
limR = c(0, 0.7)

p1 = ggplot(melt(meanRSD[reidx,reidx]*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("SD")

p2 = ggplot(melt(meanRAD[reidx,reidx]*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("AD")

p3 = ggplot(melt(meanRHC[reidx,reidx]*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC")

p4 = ggplot(melt(meanRSD_FDR05), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("FDR 0.05")

p5 = ggplot(melt(meanRAD_FDR05), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("FDR 0.05")

p6 = ggplot(melt(meanRHC_FDR05), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limR, space = "Lab", name = "R") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("FDR 0.05")

plot_grid(p1, p2, p3, p4, p5, p6, ncol = 3, nrow = 2, rel_widths = c(1, 1))
```

## Fisher z-transformation
```{r Fisher z-transformation and one-way ANOVA across groups}
Z = 0.5*(log(1+R) - log(1-R))
```


## One-way ANOVA across groups
Null hypothesis is AD = SD = HC.
```{r One-way ANOVA across groups}
Fvals = array(0, dim=c(N_ROIS, N_ROIS))
Pvals = array(NA, dim=c(N_ROIS, N_ROIS))
for (i in 1:N_ROIS) {
  for (j in 1:N_ROIS) {
    if (i != j ) {
      y = Z[i,j,]
      result = summary(aov(y[infos$included] ~ infos$group[infos$included]))
      result = result[[1]]
      Fvals[i,j] = result$`F value`[1]
      Pvals[i,j] = result$`Pr(>F)`[1]
    }
  }
}

# uncorrected
Fvals_01 = Fvals
Fvals_01[Pvals > 0.01] = 0

# FDR
Pvals_ = Pvals[reidx,reidx]
Pvals_[!tri] = NA
Pvals_FDR05 = array(p.adjust(Pvals_, method = "fdr"), dim=c(N_ROIS,N_ROIS))
Fvals_FDR05 = Fvals[reidx,reidx]
Fvals_FDR05[Pvals_FDR05 > 0.05 | is.na(Pvals_FDR05)] = 0
```

## Plot: Difference between groups
```{r Plot: Difference between groups, fig.height=2.3, fig.width=9}
limF = c(0, 12)

p1 = ggplot(melt(Fvals[reidx,reidx]*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "blue",
                       midpoint = 0, limit = limF, space = "Lab", name = "F") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("F values")

p2 = ggplot(melt(Fvals_01[reidx,reidx]*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "blue",
                       midpoint = 0, limit = limF, space = "Lab", name = "F") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("p < .01 uncor.")

p3 = ggplot(melt(Fvals_FDR05), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "blue",
                       midpoint = 0, limit = limF, space = "Lab", name = "F") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12,angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("FDR .05")

plot_grid(p1, p2, p3, ncol = 3, nrow = 1, rel_widths = c(1, 1))

```
## Significant edges from ANOVA
```{r Significant edges from ANOVA}
idx = which(t(Fvals_FDR05) > 0, arr.ind = T)
cbind(as.character(rois$AAL_Name[match(idx[,1], rois$position)]),
      as.character(rois$AAL_Name[match(idx[,2], rois$position)]))
```

## t-test SD vs HC
```{r t-test SD vs HC}
HCvsSD=list()
HCvsAD=list()

HCvsSD$Tvals = array(0, dim=c(N_ROIS, N_ROIS))
HCvsSD$Pvals = array(NA, dim=c(N_ROIS, N_ROIS))

HCvsAD$Tvals = array(0, dim=c(N_ROIS, N_ROIS))
HCvsAD$Pvals = array(NA, dim=c(N_ROIS, N_ROIS))

for (i in 1:N_ROIS) {
  for (j in 1:N_ROIS) {
    if (i != j ) {
      
      y = Z[i,j,]
      
      idx = (infos$isSD | infos$isHC) & infos$included
      result = t.test(y[idx] ~ infos$group[idx], var.equal=T,
                      alternative = "greater")
      HCvsSD$Tvals[i,j] = result[[1]]
      HCvsSD$Pvals[i,j] = result[[3]]
      
      idx = (infos$isAD | infos$isHC) & infos$included
      result = t.test(y[idx] ~ infos$group[idx], var.equal=T,
                      alternative = "less")
      HCvsAD$Tvals[i,j] = -result[[1]]
      HCvsAD$Pvals[i,j] = result[[3]]
    }
  }
}

# FDR correction
a = HCvsSD$Pvals[reidx,reidx]
b = HCvsAD$Pvals[reidx,reidx]
a[!tri] = NA
b[!tri] = NA

M = array(p.adjust(rbind(a,b), method = "fdr"), dim=c(2*N_ROIS, N_ROIS))
HCvsSD$Pvals_FDR05 = M[1:14,]
HCvsAD$Pvals_FDR05 = M[15:28,]

HCvsSD$Tvals_FDR05 = HCvsSD$Tvals[reidx,reidx]
HCvsAD$Tvals_FDR05 = HCvsAD$Tvals[reidx,reidx]

HCvsSD$Tvals_FDR05[HCvsSD$Pvals_FDR05 > 0.05 | is.na(HCvsSD$Pvals_FDR05)] = 0
HCvsAD$Tvals_FDR05[HCvsAD$Pvals_FDR05 > 0.05 | is.na(HCvsAD$Pvals_FDR05)] = 0
```

## Plot: t-Test
```{r Plot: t-Test, fig.height=4.5, fig.width=6}
limF = c(-3, 5)

p1 = ggplot(melt(HCvsSD$Tvals[reidx,reidx]*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limF, space = "Lab", name = "t") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > SD")

p2 = ggplot(melt(HCvsAD$Tvals[reidx,reidx]*tri), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limF, space = "Lab", name = "t") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > AD")

p3 = ggplot(melt(HCvsSD$Tvals_FDR05), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limF, space = "Lab", name = "t") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > SD (FDR 0.05)")

p4 = ggplot(melt(HCvsAD$Tvals_FDR05), aes_string(x = "Var1", y = "Var2", fill = "value")) +
  geom_tile(color = "gray60") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limit = limF, space = "Lab", name = "t") +
  theme(axis.ticks = element_blank(), axis.line = element_blank(),
        text = element_text(size=12), plot.title = element_text(size=12),
        axis.text.x = element_text(size=12, angle=0),
        axis.text.y = element_text(size=12)) + 
  scale_y_reverse() + ggtitle("HC > AD (FDR 0.05)")

plot_grid(p1, p2, p3, p4, ncol = 2, nrow = 2, rel_widths = c(1, 1))
```

## Significant edges from t-test
HC > SD
```{r Significant edges from t.test}
idx = which(t(HCvsSD$Tvals_FDR05) > 0, arr.ind = T)
cbind(as.character(rois$AAL_Name[match(idx[,1], rois$position)]),
      as.character(rois$AAL_Name[match(idx[,2], rois$position)]))
```

HC > AD
```{r Significant edges from t.test cont.}
idx = which(t(HCvsAD$Tvals_FDR05) > 0, arr.ind = T)
cbind(as.character(rois$AAL_Name[match(idx[,1], rois$position)]),
      as.character(rois$AAL_Name[match(idx[,2], rois$position)]))
```


## Group difference
```{r Group difference, fig.height=2.3, fig.width=5}
d = data.frame(R = R[3,1,], Z = Z[3,1,], group=infos$group)
d = subset(d, infos$included)

p1 = ggplot(d, aes(x=group, y=R)) + geom_violin() + geom_boxplot(width=0.2) +
  geom_point(shape=1, color="gray70", size=4, position = position_jitter(width = 0.2, height = 0.0002)) + 
  ggtitle("Pearson's R")

p2 = ggplot(d, aes(x=group, y=Z)) + geom_violin() + geom_boxplot(width=0.2) +
  geom_point(shape=1, color="gray70", size=4, position = position_jitter(width = 0.2, height = 0.0002)) + 
  ggtitle("Fisher's Z")

plot_grid(p1, p2, ncol = 2, nrow = 1, rel_widths = c(1, 1))

summary(aov(R ~ group, data=d))
summary(aov(Z ~ group, data=d))

```


